### 2021-12-19 Manuscript.Rmd
updated deep root qn

For each gene tree, we generated summary statistics of the branch lengths in the tree (maximum, minimum, standard deviation) as well as the fraction of branches that have a default length value of `r default_length_val` (that is indicative of branch length=0). The majority of gene trees have branches with a maximum length around 1. Gene trees were filtered to exclude trees with branches whose maximum length is >2 and that had more than 0.25 branches with the default length value, representing `r (length(exclude_hashes)/raw_gene_trees_n) %>% round(4) *100`% of gene trees. The goal of this is to exclude trees that include very long branches, or a large proportion of branches with very short branch lengths. Gene tree nodes were annotated as speciation events or duplication events, based on assignment by Orthofinder 2 [@emms2019orthofinder]. Speciation nodes were subsequently assigned a node ID equivalent to the species tree node, using species tip names from the gene tree to determine the most common recent ancestor in the species tree, using the `phytools` package. Due to the use of the species-overlap method by Orthofinder 2, some clades of single copy genes were assigned  as speciation events, although the topology is inconsistent with the species tree. To avoid time calibration issues, due to descendant nodes being assigned the same species node ID, descendant speciation nodes with the same species node ID were marked as null, and gene trees with nodes greater than 0.3 null nodes per internal node were excluded -- indicating widespread topological differences with the species tree. Additionally, only trees with one or more speciation events were retained, as speciation events are used for time calibrations. The gene trees were then time calibrated to the species tree using `chronos()` in the `ape` package, so that the branch lengths were scaled to the same equivalent length across all gene trees [@ape2004]. Some gene trees could not be calibrated against the node constraints from the species tree and were discarded. Additionally we calculated the maximum node age for each gene tree, and excluded gene trees with roots that exceeded a maximum root depth of 5, which can be indicative of calibration problems. Tips without expression values were then pruned out of the tree. Gene trees with fewer than three expression values at the tips were discarded, retaining only trees with three or more values.  Pruned and unpruned calibrated gene tree files are available as supplementary data.

>>>>

For each gene tree, we generated summary statistics of the branch lengths in the tree (maximum, minimum, standard deviation) as well as the fraction of branches that have a default length value of `r default_length_val` (that is indicative of branch length=0). The majority of gene trees have branches with a maximum length around 1. Gene trees were filtered to exclude trees with branches whose maximum length is >2 and that had more than 0.25 branches with the default length value, representing `r (length(exclude_hashes)/raw_gene_trees_n) %>% round(4) *100`% of gene trees. The goal of this is to exclude trees that include very long branches, or a large proportion of branches with very short branch lengths. Gene tree nodes were annotated as speciation events or duplication events, based on assignment by Orthofinder 2 [@emms2019orthofinder]. Speciation nodes were subsequently assigned a node ID equivalent to the species tree node, using species tip names from the gene tree to determine the most common recent ancestor in the species tree, using the `phytools` package. Due to the use of the species-overlap method by Orthofinder 2, some clades of single copy genes were assigned  as speciation events, although the topology is inconsistent with the species tree. To avoid time calibration issues, due to descendant nodes being assigned the same species node ID, descendant speciation nodes with the same species node ID were marked as null, and gene trees with nodes greater than 0.3 null nodes per internal node were excluded -- indicating widespread topological differences with the species tree. Additionally, only trees with one or more speciation events were retained, as speciation events are used for time calibrations. The gene trees were then time calibrated to the species tree using `chronos()` in the `ape` package, so that the branch lengths were scaled to the same equivalent length across all gene trees [@ape2004]. Some gene trees could not be calibrated against the node constraints from the species tree and were discarded. Additionally we calculated the maximum node age for each gene tree, and excluded gene trees with roots that exceeded a maximum root depth of 5, which can be indicative of calibration problems (the root of the species tree is 1). Only a single tree was excluded as a result of a very deep root. Tips without expression values were then pruned out of the tree. Gene trees with fewer than three expression values at the tips were discarded, retaining only trees with three or more values.  Pruned and unpruned calibrated gene tree files are available as supplementary data.
